When to Compete and When to Cooperate: Analysis on Strategies of Services within Communities

### 2012-11-28 papers/coopv2/ICSOC Submission/ICSOC 2012.tex
babak changes

% first the title is needed
\title{Analyzing Coopetition Strategies of Services within Communities}

@@

\author{Babak Khosravifar$^1$
\and Mahsa Alishahi$^1$ \and Ehsan Khosrowshahi Asl$^1$ \and Jamal
Bentahar$^1$\and Rabeb Mizouni$^2$\and Hadi Otrok$^2$}

@@

\institute{$^1$Concordia University, Canada, $^2$Khalifa University, UAE\\
$b\_khosr@encs.concordia.ca$, $m\_ali@encs.concordia.ca$, $e\_khosr@encs.concordia.ca$,\\$bentahar@ciise.concordia.ca$,
$rabeb.mizouni@kustar.ac.ae$,$hadi.otrak@kustar.ac.ae$}

@@

Figure \ref{architectureFigure} illustrates the architecture of a
typical community aggregating a number of web services with
different interactive strategies. Some of them compete for the
task where they directly deal with the master. Some others
cooperate in the associated task where they only deal with the
competed web service as the task leader and do not directly
interact with the master (the master deals only with the web
service that has bid for the task, which is responsible of
choosing its collaborative network). In both sets, some web
services are for certain moments out of any collaboration network.
We highlight details of the interactive strategies in the rest of
this section.
%Figure \ref{algorithm1} %exposes the details of task allocation mechanism that is used to select the web services.
In the proposed system, the master sorts the competing web
services based on some parameters (such as reputation) that we
explain in the rest of this section. If the web service is busy or
unwilling to take the task, the master allocates the task to the
second competing web service in the list. There is a chance that
some tasks could not be assigned to any web service. These tasks
are accumulated to the task pool to be allocated in the next task
allocation round. Upon allocation of the task, the web service is
responsible to offer the required QoS that is stated in the task
being generated by a consumer. Afterwards,  the master rewards or
penalizes the competing web service by upgrading or degrading its
reputation according to the offered QoS compared with the required
one. This comparison influences the sorting mechanism used by the
master to allocate the tasks in further task allocation rounds.

@@

\begin{figure}[h]
\centering
\includegraphics[scale=0.45]{decision1++.eps}
\caption{Decision making process over competitive and cooperative
strategies.} \label{decisionMakingProcedure}
\end{figure}

@@

In this section, we provide an empirical analysis over the
observed results regarding the characteristics of intelligent web
services hosted in different communities of web services. In the
implemented system, we simulate the behaviors of service consumers
as request generators, web services as service providers, and
master web services as community representatives. These entities
are developed with respect to what is explained in Section
\ref{Preliminaries}. The objective is to investigate the
effectiveness of the proposed strategic system on intelligent
web services' overall budge and also the average quality and quantity of tasks performed by 
community web services which directly affects user satisfaction. 
%Before conducting the simulation, the
%expected behavior of our model is to enhance the budget of web
%services when using our proposed strategic interacting system.
%Although this result is promising, we would like to investigate
%the effectiveness of the proposed strategic system on intelligent
%web services' overall budget compared to the best case. Moreover,
We study the overall performance of the community hosting the
reasoning-empowered web services compared to the ones hosting
stochastic and purely competitive services. The simulation
application is written in \textit{C\#} using \textit{Visual Studio}. 
Developed web services are initialized with values taken
from a real dataset that includes $2507$ real web services
functioning on the web. The dataset records the QoS values of $9$
parameters including \textit{availability}, \textit{throughput},
and
\textit{reliability} \cite{Almasri}.  %These QoS values
%are determined by monitoring the web services over a $6$-day
%period.

We start our discussions with cumulative budget comparison
regarding different communities within which services follow
different reasoning techniques. Figure \ref{Graph1} part (a)
illustrates three graphs for three different communities. Each
community hosts web services that follow different reasoning
techniques: (1) a community that follows the interactive reasoning
techniques presented in this paper (referred to as coopetitive);
(2) a community that follows a random reasoning technique so
decisions about selecting competitive or cooperative strategies
are totally random (referred to as random coopetitive); and (3) a
competitive community where all services follow the competitive
strategy (referred to as competitive). The proposed model's
reasoning mechanism enables services to effectively select their
interacting strategies and the obtained budget represents the best
outcome over the strategic decision making procedure they run all
the time. This procedure avoids cases where a service selects the
competitive strategy but gets refused to obtain a task from the
master. The procedure allows services to make decisions that
maximize their utilities, so that if the web service cannot
compete, the procedure would suggest to collaborate, which is
better than competing and failing to obtain the task. In this case
(i.e., competing and not getting the task), the service stays idle
but still pays the community membership fee, which means losing
utility. The developed strategic decision making mechanism leads
some web services to follow cooperative strategies that overall
maintain an optimal community budget. In the same figure, we
observe the cumulative budget of a community where services follow
random interacting strategies. The outcome is clearly lower
because services at each run randomly decide over their acting
strategies. This potentially influences the community budget
because a low quality service if randomly selects to follow the
competitive strategy, it will fail to perform a task with high QoS requirements. This kind of
strategy selection is totally stochastic while the task allocation
algorithm follows a logical path. The ideal system is the one that
analyzes the optimal strategic path and consistently follows
strategies that bring maximum outcome. The result regarding the
community that follows the random strategy shows how stochastic
decision making degrades the community budget, but still this
result outperforms the one of the purely competitive community. In
the last case, all the services follow the competitive strategy
and the frequency of getting rejected from a task is relatively
high. In a community with limited number of tasks, the competitive
strategy for all services highly influences the community budget
because a potential group of services are losing at every run.

@@

In figure \ref{membership_fee_graph} we analyze the scenario where membership fee is high which is analogous to situations where task income is low. For any reason our web services may need to stay in a particular community with economic downtimes. We need verify how well the web services can perform in these situations. \ref{membership_fee_graph} part (a) depicts the results in normal scenario. In part (b) we have increased the membership fee to the high amount of 5 units per run. Web services will need to pay 5 units on each iteration of tasks being offered. As results the strategic coopetitive decision making process helps web services to still have benefit and random or all compete methods will lose a lot in these scenarios since making mistakes and trying to compete for jobs they cannot afford while having to pay higher membership fees is making them lose budget in a faster pace. In part \ref{membership_fee_graph}(c) the membership is higher than average of possible income of what tasks offer. It might not be rational for selfish self-interest web services to stay in these communities from utility point of view but as mentioned there might be other political reasons they might need to stay in these communities. As results show the random and all compete strategies make web services bank deplete in no time. However our strategic coopetitive decision making process loses budget in a slower pace.

@@

The web services are selfish, self interested agents. They try to maximise their own utiliy. Now we analyse how their strategy effects the social welfare or from user and community point of view how good the tasks are being performed. This directly affects user satisfation and community reputation in general. The Higher  quality and quantity of tasks perfromed leads to higher user satisfaction for the community which results in better reputaiton for the community. The results in figure \ref{graph_task} shows the quality of quantity of tasks being done successfuly in three communities adopting the three different mentioned strategy decision algorithms. Figure \ref{graph_task} Part (a) and (b) depicts the number tasks succesfuly done. When web services are not capable of performing tasks alone and decide to compete for tasks with high QoS requirements they usually fail to perform the task how ever if they opt for cooperating although they have less income, they have higher chance of performing tasks collabratively with other web serives. Figure \ref{graph_task}Part (b) shows the probablilty of successfully performin tasks. As results suggest web services using our coopetitive strategy when deciding whether to compete or cooperate have higher chance of tasks to be done than random or all compete strategies. The quality of tasks performed are depicted in Figure \ref{graph_task} Part (c). It shows the average QoS of tasks perfromed which are finished and done sucessfuly. Colabrating increases the performed QoS of tasks, therefore forming collabrative groups improves user satisfaction. As results suggest, the coopetetive community outperforms the random and all compete communities in quality of tasks performed.

>>>>

% first the title is needed
\title{When to Compete and When to Cooperate: Analysis on Strategies of Services within Communities}

@@

\author{Babak Khosravifar$^1$\and Ehsan Khosrowshahi Asl$^2$
\and Jamal Bentahar$^2$\and Rabeb Mizouni$^3$\and Hadi Otrok$^3$}

@@

\institute{$^1$McGill University, Canada, $^2$ Concordia University, Canada, $^3$Khalifa University, UAE\\
$babak.khosravifar@mcgill.ca$, $e\_khosr@encs.concordia.ca$, $bentahar@ciise.concordia.ca$,
$rabeb.mizouni@kustar.ac.ae$,$hadi.otrak@kustar.ac.ae$}

@@

Figure \ref{architectureFigure} illustrates the architecture of a
typical community aggregating a number of web services with
different interactive strategies. Some of them compete for the
task where they directly deal with the master. Some others
cooperate in the associated task where they only deal with the
competed web service as the task leader and do not directly
interact with the master (the master deals only with the web
service that has bid for the task, which is responsible of
choosing its collaborative network). In both sets, some web
services are for certain moments out of any collaboration network.
We highlight details of the interactive strategies in the rest of
this Section.
%Figure \ref{algorithm1} %exposes the details of task allocation mechanism that is used to select the web services.
In the proposed system, the master sorts the competing web
services based on some parameters (such as reputation) that we
explain in the rest of this section. If the web service is busy or
unwilling to take the task, the master allocates the task to the
second competing web service in the list. There is a chance that
some tasks could not be assigned to any web service. These tasks
are accumulated to the task pool to be allocated in the next task
allocation round. Upon allocation of the task, the web service is
responsible to offer the required QoS that is stated in the task
being generated by a consumer. Afterwards,  the master rewards or
penalizes the competing web service by upgrading or degrading its
reputation according to the offered QoS compared with the required
one. This comparison influences the sorting mechanism used by the
master to allocate the tasks in further task allocation rounds.

@@

\begin{figure}[h]
\centering
\includegraphics[scale=0.55]{decision1++.eps}
\caption{Decision making process over competitive and cooperative
strategies.} \label{decisionMakingProcedure}
\end{figure}

@@

In this section, we provide an empirical analysis over the
observed results regarding the characteristics of intelligent web
services hosted in different communities of web services. In the
implemented system, we simulate the behaviors of service consumers
as request generators, web services as service providers, and
master web services as community representatives. These entities
are developed with respect to what is explained in Section
\ref{Preliminaries}. In this Section, the objective is to investigate the
effectiveness of the proposed strategic system on intelligent
web services' overall budget and also the average quality and quantity of tasks performed by 
community web services which directly affects user satisfaction. 
%Before conducting the simulation, the
%expected behavior of our model is to enhance the budget of web
%services when using our proposed strategic interacting system.
%Although this result is promising, we would like to investigate
%the effectiveness of the proposed strategic system on intelligent
%web services' overall budget compared to the best case. Moreover,
To verify these objectives, we study the overall performance of the community hosting the
reasoning-empowered web services compared to the ones hosting
stochastic and purely competitive services. The simulation
application is written in \textit{C\#} using \textit{Visual Studio}. 
Developed web services are initialized with values taken
from a real dataset that includes $2507$ real web services
functioning on the web. The dataset records the QoS values of $9$
parameters including \textit{availability}, \textit{throughput},
and
\textit{reliability} \cite{Almasri}.  %These QoS values
%are determined by monitoring the web services over a $6$-day
%period.

We start our discussions with cumulative budget comparison
regarding different communities within which services follow
different reasoning techniques. Figure \ref{Graph1} part (a)
illustrates three graphs for three different communities. Each
community hosts web services that follow different reasoning
techniques: (1) a community that follows the interactive reasoning
techniques presented in this paper (referred to as coopetitive);
(2) a community that follows a random reasoning technique so
decisions about selecting competitive or cooperative strategies
are totally random (referred to as random coopetitive); and (3) a
competitive community where all services follow the competitive
strategy (referred to as competitive). The proposed model's
reasoning mechanism enables services to effectively select their
interacting strategies and the obtained budget represents the best
outcome over the strategic decision making procedure they run all
the time. This procedure avoids cases where a service selects the
competitive strategy but gets refused to obtain a task from the
master. The procedure allows services to make decisions that
maximize their utilities, so that if the web service cannot
compete, the procedure would suggest to collaborate, which is
better than competing and failing to obtain the task. In this case
(i.e., competing and not getting the task), the service stays idle
but still pays the community membership fee, which means losing
utility. The developed strategic decision making mechanism leads
some web services to follow cooperative strategies that overall
maintain an optimal community budget. In the same Figure, we
observe the cumulative budget of a community where services follow
random interacting strategies. The outcome is clearly lower
because services at each run randomly decide over their acting
strategies. This potentially influences the community budget
because a low quality service if randomly selects to follow the
competitive strategy, it will fail to perform a task with high QoS requirements. This kind of
strategy selection is totally stochastic while the task allocation
algorithm follows a logical path. The ideal system is the one that
analyses the optimal strategic path and consistently follows
strategies that bring maximum outcome. The result regarding the
community that follows the random strategy shows how stochastic
decision making degrades the community budget, but still this
result outperforms the one of the purely competitive community. In
the last case, all the services follow the competitive strategy
and the frequency of getting rejected from a task is relatively
high. In a community with limited number of tasks, the competitive
strategy for all services highly influences the community budget
because a potential group of services are losing at every run.

@@

In Figure \ref{membership_fee_graph} we analyze the scenario where membership fee is high which is analogous to situations where task income is low. For any reason our web services may need to stay in a particular community with economic downtimes. We need verify how well the web services can perform in these situations. \ref{membership_fee_graph} part (a) depicts the results in normal scenario. In part (b) we have increased the membership fee to the high amount of 5 units per run. Web services will need to pay $5$ units on each iteration of tasks being offered. As results the strategic coopetitive decision making process helps web services to still have benefit and random or all compete methods will lose a lot in these scenarios since making mistakes and trying to compete for jobs they cannot afford while having to pay higher membership fees is making them lose budget in a faster pace. In part \ref{membership_fee_graph}(c) the membership is higher than average of possible income of what tasks offer. It might not be rational for selfish self-interest web services to stay in these communities from utility point of view but as mentioned there might be other political reasons they might need to stay in these communities. As results show the random and all compete strategies make web services bank deplete in no time. However our strategic coopetitive decision making process loses budget in a slower pace.

@@

The web services are selfish, self interested agents. They try to maximise their own utiliy. Now we analyse how their strategy effects the social welfare or from user and community point of view how good the tasks are being performed. This directly affects user satisfation and community reputation in general. The Higher  quality and quantity of tasks perfromed leads to higher user satisfaction for the community which results in better reputaiton for the community. The results in Figure \ref{graph_task} shows the quality of quantity of tasks being done successfuly in three communities adopting the three different mentioned strategy decision algorithms. Figure \ref{graph_task} Part (a) and (b) depicts the number tasks succesfuly done. When web services are not capable of performing tasks alone and decide to compete for tasks with high QoS requirements they usually fail to perform the task how ever if they opt for cooperating although they have less income, they have higher chance of performing tasks collabratively with other web serives. Figure \ref{graph_task}Part (b) shows the probablilty of successfully performin tasks. As results suggest web services using our coopetitive strategy when deciding whether to compete or cooperate have higher chance of tasks to be done than random or all compete strategies. The quality of tasks performed are depicted in Figure \ref{graph_task} Part (c). It shows the average QoS of tasks perfromed which are finished and done sucessfuly. Colabrating increases the performed QoS of tasks, therefore forming collabrative groups improves user satisfaction. As results suggest, the coopetetive community outperforms the random and all compete communities in quality of tasks performed.
